---
title: "Bikeshare Data"
author: "Nate Smullen"
date: "10/27/2022"
output: html_notebook
---

Bike Share Data

##Load Packages

```{r}
library(tidyverse)
library(here)
library(janitor)
library(maps)
library(lubridate)
library(gbfs)

library(taskscheduleR) #might be useful for running code every minute
```

##Load Bikeshare Data

```{r}
thisfile=here("data_raw","202209-capitalbikeshare-tripdata.csv")

bike = read_csv(thisfile) %>% clean_names() %>%
  na.omit(bike)
```

```{r}
DC <- map_data("state") %>% filter(region=="district of columbia")
```

```{r}
ggplot(bike, mapping = aes(x=start_lng, y=start_lat)) +
  geom_polygon(data = DC, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_point(alpha = .01)
```

```{r}
bikea = bike %>% mutate(duration = as.numeric((ended_at - started_at)/60),
                        hour_of_day=hour(started_at),
                        day_of_week=wday(started_at,label = T))
```

```{r}
bikea %>% filter(duration < 100, duration > 0) %>%
  ggplot(mapping = aes(duration)) +
         geom_histogram(bins = 300) + facet_wrap(~day_of_week)
```

```{r}
bikeb = bikea %>% filter(duration >0)
```

```{r}
bikec = bikeb %>% filter(mday(started_at) == 3) %>% slice_sample(n = 100)
```

```{r}
bikec %>% select(start = started_at, end = ended_at)
```
Make it a long dataset
```{r}
bikec %>% select(start = started_at, end = ended_at) %>%
  pivot_longer(start:end)
```

pick namges and order by time

```{r}
bikec %>% select(start = started_at, end = ended_at) %>%
  pivot_longer(start:end, names_to = "type", values_to = "time") %>% 
  arrange(time)
```
add counter
```{r}
bikec %>% select(start = started_at, end = ended_at) %>%
  pivot_longer(start:end, names_to = "type", values_to = "time") %>% 
  arrange(time) %>% 
  mutate(increment = case_when(
    type == "start" ~ 1,
    type == "end" ~ -1
  ))
```
cumulative sum for that column
```{r}
bikec %>% select(start = started_at, end = ended_at) %>%
  pivot_longer(start:end, names_to = "type", values_to = "time") %>% 
  arrange(time) %>% 
  mutate(increment = case_when(
    type == "start" ~ 1,
    type == "end" ~ -1
  )) %>%
  mutate(riders = cumsum(increment))
```
```{r}
bikec %>% select(start = started_at, end = ended_at) %>%
  pivot_longer(start:end, names_to = "type", values_to = "time") %>% 
  arrange(time) %>% 
  mutate(increment = case_when(
    type == "start" ~ 1,
    type == "end" ~ -1
  )) %>%
  mutate(riders = cumsum(increment)) %>%
  ggplot(aes(time, riders)) + geom_step()
```
Do it with the whole dataset at september 3
```{r}
bikeb %>% filter(mday(started_at) == 3) %>%
  select(start = started_at, end = ended_at) %>%
  pivot_longer(start:end, names_to = "type", values_to = "time") %>% 
  arrange(time) %>% 
  mutate(increment = case_when(
    type == "start" ~ 1,
    type == "end" ~ -1
  )) %>%
  mutate(riders = cumsum(increment)) %>%
  ggplot(aes(time, riders)) + geom_step() +
  scale_x_datetime(limits = as_datetime(c("2022-09-03", "2022-09-04")))
```
Facet the daily graphs for the whole month to see the whole dataset
```{r}
bikeb %>% filter(month(started_at) == 9, month(ended_at) == 9) %>%
  select(start = started_at, end = ended_at) %>%
  pivot_longer(start:end, names_to = "type", values_to = "time") %>% 
  arrange(time) %>% 
  mutate(increment = case_when(
    type == "start" ~ 1,
    type == "end" ~ -1
  )) %>%
  mutate(riders = cumsum(increment)) %>%
  ggplot(aes(time, riders)) + geom_step() +
  facet_wrap(~mday(time), scales = "free_x", ncol = 7)
```
*During the weekdays, there are 2 peaks in number of riders, vs weekends having
  one peak
*For weekdays, there are more people riding during the afternoon peak than
  during the morning peak
*September 5 is a weekday that only 1 peak, but it was labor day, so many people
  did not have to go to work/school.
*Fridays typically increases throughout the day, making the the first peak less
  defined
*There was a storm that hit DC on the 11th and 12th, likely causing the decrease
  in number of riders that day
*The 30th had some rain in the evening, perhaps explaining the decrease in the
  number of riders for the second peak


Compare casual riders to members
```{r}
bikeb %>%
  filter(month(started_at)==9,month(ended_at)==9) %>%
select(member_casual,start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="time") %>% arrange(time) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) %>%
  group_by(member_casual) %>%
  mutate(riders=cumsum(increment)) %>%
    ggplot(aes(time,riders,color=member_casual)) + geom_step() +
  facet_wrap(~mday(time),scales = "free_x",ncol = 7)
```
```{r}
bikeb %>% filter(duration < 100) %>%
  group_by(member_casual) %>%
  ggplot(aes(x = duration, color = member_casual)) + geom_histogram(bins = 300) +
  facet_wrap(~day_of_week)
```

```{r fig.height = 10}
bikeb %>% filter(duration < 100) %>%
  ggplot(aes(x = duration)) + geom_histogram(bins = 300) +
  facet_wrap(~hour(started_at), scales = 'free_y', ncol = 6)
```

```{r}
df_cities = get_gbfs_cities()
```


```{r}
station_info <- get_station_information("cabi")
free_bikes <- get_free_bike_status("cabi")
station_status <- get_station_status("cabi")
```
Next Question:
How close does it get to no bikes being available in some locations?

See live data, run code with a timer, pick 3 or 4 that are close to running out
Check every minute during rush hours

```{r}
station_status <- get_station_status("cabi") %>% 
  mutate(total_bikes_available = 
  num_bikes_available + num_docks_available + num_ebikes_available)
```

```{r}
free_bikes %>% ggplot(aes(x = lon, y = lat)) +  
  geom_polygon(data = DC, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_point()
```


```{r}
#code that runs every minute for an hour
#problem with this code is that it doesn't account for the time it takes to do 
#The calculation
i = 0
while(i < 60) {
  print(paste("its been", i, "minutes since this started, give or take a bit"))
  i = i + 1
  
  Sys.sleep(60)
}
```

