---
title: "Lab 4: Spatial Data with Denny's and La Quinta"
author: "Nate Smullen"
date: "10/16/2022"
output: html_notebook
---

Visualizing spatial data

##Load Packages

```{r}
library(tidyverse)
library(dsbox)
```

##Load States Data

```{r}
states <- read_csv("states.csv")
```

1.
```{r}
nrow(dennys)
ncol(dennys)
names(dennys)
```
There are 1643 rows in the Denny's dataset, each representing a different 
Denny's restaurant. There are 6 columns in the Denny's dataset, each 
representing a different attribute. The attributes are address, city, state, zip
code, longitude, and latitude.

2.
```{r}
nrow(laquinta)
ncol(laquinta)
names(laquinta)
```
There are 909 rows in the LaQuinta's dataset, each representing a different 
LaQuinta's motel. There are 6 columns in the LaQuinta's dataset, each 
representing a different attribute. The attributes are address, city, state, zip
code, longitude, and latitude.

3. There are several LaQuinta's locations outside of the US. There are some in 
Canada, Mexico, China, New Zealand, Turkey, United Arab Emirates, Chile, and 
Columbia. There are no Denny's locations listed on the website that are outside
of the US.

4. The locations outside of the US still have a value in the state column, so 
you cannot filter by null values in that column. You could check to make sure
each state is actually a state in the US, however the state abbreviation for a
location in a different country may be the same as one in the US. Instead,
you could try to filter by zip codes that are in the US. Zip codes, however, are
not unique globally, so you could run into the issue where a foreign zip code
matches one in the US. Therefore, the safest way to determine if a state is in
the US is to filter based on combinations of longitudes and latitudes that are
inside of the US.

5.
```{r}
dennys %>%
  filter(!(state %in% states$abbreviation))
```
The Denny's dataset filtered by states not in the US is empty, meaning there
are no locations outside of the US.

6.
```{r}
dennys <- dennys %>%
  mutate(country = "United States")
```

7.
```{r}
laquinta %>%
  filter(!(state %in% states$abbreviation))
```
* ON, BC are in Canada
* ANT is in Colombia
* Ag, QR, Ch, NL, VE, PU, SL are in Mexico
* FM is in Honduras

8.
```{r}
laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state %in% c("AG" ,"QR","CH","NL","VE","PU","SL") ~ "Mexico",
    state == "FM"            ~ "Honduras"))
```
```{r}
laquinta <- laquinta %>%
  filter(country == "United States")
```

9.
```{r}
dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))
```

10.
```{r}
dennys <- dennys %>%
  mutate(establishment = "Denny's")
laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")
```
```{r}
dn_lq <- bind_rows(dennys, laquinta)
```
```{r}
ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point(alpha = .5)
```
11.
```{r}
dn_lq_nc = dn_lq %>% filter(state == "NC")

ggplot(dn_lq_nc, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point(alpha = .5)
```
Visually, Mitch Hedberg's joke does not hold in North Carolina


12.
```{r}
dn_lq_nc = dn_lq %>% filter(state == "TX")

ggplot(dn_lq_nc, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point(alpha = .5)
```
Visually, Mitch Hedberg's joke holds in Texas.







