---
title: "MA131 and MA132 Enrollment Prediction"
author: "Nate Smullen"
date: "10/3/2022"
output: html_notebook
---

Class play with MA131 and MA132 enrollment data.

##Load Packages

```{r}
library(tidyverse)
library(here)
library(janitor)
```

##Load Enrollment Data

```{r}
thisfile=here("data_raw","CU_ACAD_DEPT_WHEN_OFF_ENRL_22740.csv")

df1=read_csv(thisfile) %>% clean_names()

```

```{r}
df2 <- aggregate(df1$tot_enrl, list(df1$term_8, df1$title), sum)
colnames(df2) <- c("term", "course_title", "tot_students_enrolled")

df2 %>% ggplot(aes(fill=course_title, x=term, y=tot_students_enrolled)) + 
    geom_bar(position="dodge", stat="identity")

```

Above plot has the correct labels but it is not sorted by time because of the
alphabetical sort. We do not know how to order it correctly, but the labels
on the X axis are what we want.

```{r}
df2 <- aggregate(df1$tot_enrl, list(df1$term_7, df1$title), sum)
colnames(df2) <- c("term", "course_title", "tot_students_enrolled")

df2 %>% ggplot(aes(fill=course_title, x=term, y=tot_students_enrolled)) + 
    geom_bar(position="dodge", stat="identity")

```

Above plot has the correct order, as it is sorted from oldest to newest, however
it is very difficult to read which section is which. Ideally, a mix between
this plot and the previous plot is what we are aiming for, but we do not yet
know how to do it.