---
title: "NYC Flights"
author: "Nate Smullen"
date: "2022-09-09"
output: html_notebook
---

Class play with NYC flights data.

## Load Packages

```{r}
library(nycflights13)
library(tidyverse)
library(here)
library(janitor)
```

## Load flight data

```{r}
df1 = flights
```

## Class work

```{r}
df2 = df1 %>%
  filter(origin == "MIA", month == 9)

df3 = df1 %>%
  filter(month == 9, dest == "MIA")

df4 = df1 %>%
  filter(month == 1, dest == "MIA")

df5 = df1 %>%
  filter(month == 7, dest == "MIA")

df6 = df1 %>%
  filter(dest == "MIA")

df6 %>% ggplot(aes(month)) + geom_bar()
```
Analyzing the graph above we note:

* December has the most flights, perhaps due to holiday.
* February is low, perhaps due to fewer days in February.
* Flight numbers seem to correlate with days in the month.

```{r}
df7 = df1 %>%
  filter(dest %in% c("LAX", "SAN", "SFO", "ATL", "ORD", "MIA"))

df7 %>% ggplot(aes(x = month, fill = dest)) + geom_bar()
```
```{r}
df7 %>% ggplot(aes(x = month, fill = dest)) + geom_bar() + facet_wrap(~dest)
```
Observations from the above graph:

* San Diego is much smaller than the others.
* MIA is _unusually_ flat.
* ORD is summer intensive (more traffic over summer).
* ORD, LAX, SFO patterns similar: low in winter, high in summer. 
* ATL is flat, perhaps it's a hub.
* Feb is low (fewer days, winter).

